---
title: "Construction of NicheNet's ligand-target model using Omnipath Interactions"
author: "Alberto Valdeolivas: alberto.valdeolivas@bioquant.uni-heidelberg.de; Date:"
date: "29/02/2020"
output: github_document
# license: GNU-GLPv3, please check http://www.gnu.org/licenses/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, warning = FALSE, 
    message = FALSE)
```

## Abstract

This vignette shows how to build the ligand-target prior regulatory potential 
scores based on the NicheNet method but using Omnipath interactions.  

## Construct a ligand-target model from Omnipath ligand-receptor, signaling and gene regulatory networks

We first load the required packages and the networks generated on the previous
scripts: Omnipath ligand-receptor, Omnipath signaling and Omnipath (Dorothea) 
gene regulatory network. These networks are used to construct the ligand-target
model. 

```{r}
library(nichenetr)
library(tidyverse)

## We load the networks generated in the ParameterOptimization script
lr_network_Omnipath = readRDS("OmniNetworks_NNformat/lr_Network_Omnipath.rds")
sig_network_Omnipath = readRDS("OmniNetworks_NNformat/sig_Network_Omnipath.rds")
gr_network_Omnipath = readRDS("OmniNetworks_NNformat/gr_Network_Omnipath.rds")
```

## Construct NicheNet’s ligand-target model from unoptimized data source weights

The interactions available in Omnipath are well curated, and we therefore first
create a model considering the same weight for all the databases. On the other
hand, we use the optimized parameters for the PageRank algorithm. 

```{r}
## The interactions are weighted only based in the number of data sources that
## report them
All_sources <- unique(c(lr_network_Omnipath$source,
    sig_network_Omnipath$source, gr_network_Omnipath$source))

my_source_weights_df <- 
     tibble(source = All_sources, weight = rep(1,length(All_sources)))

weighted_networks <- construct_weighted_networks(
    lr_network = lr_network_Omnipath, 
    sig_network = sig_network_Omnipath, 
    gr_network = gr_network_Omnipath, 
    source_weights_df = my_source_weights_df)

## We read the results of the optimization 
resultsOptimization <- readRDS("Results/Optimization_results.rds")

optimized_parameters = resultsOptimization %>% 
    process_mlrmbo_nichenet_optimization(
        source_names = my_source_weights_df$source %>% unique())

weighted_networks <- apply_hub_corrections(
    weighted_networks = weighted_networks, 
    lr_sig_hub = optimized_parameters$lr_sig_hub, 
    gr_hub = optimized_parameters$gr_hub)
saveRDS(weighted_networks,"Results/weighted_networksNonSourceWeights.rds")
```

We now generate the matrix containing the ligand-target regulatory potential 
scores based on the weighted integrated networks. 

```{r}
ligands <- as.list(unique(lr_network_Omnipath$from))

ligand_target_matrix <- construct_ligand_target_matrix(
    weighted_networks = weighted_networks, 
    ligands = ligands, 
    algorithm = "PPR", 
    damping_factor = optimized_parameters$damping_factor, 
    ltf_cutoff = optimized_parameters$ltf_cutoff)
saveRDS(ligand_target_matrix,"Results/ligand_target_matrixNoweights.rds")
```

## Construct NicheNet’s ligand-target model from optimized data source weights

Now, we create an alternative model using the optimized weights for the 
different sources of data.

```{r}
## Here we also take into account the optimized source weights
weighted_networks <- construct_weighted_networks(
    lr_network = lr_network_Omnipath, 
    sig_network = sig_network_Omnipath, 
    gr_network = gr_network_Omnipath,
    source_weights_df = optimized_parameters$source_weight_df)

weighted_networks <- apply_hub_corrections(
    weighted_networks = weighted_networks, 
    lr_sig_hub = optimized_parameters$lr_sig_hub, 
    gr_hub = optimized_parameters$gr_hub)

ligand_target_matrix = construct_ligand_target_matrix(
    weighted_networks = weighted_networks, 
    ligands = ligands, 
    algorithm = "PPR", 
    damping_factor = optimized_parameters$damping_factor, 
    ltf_cutoff = optimized_parameters$ltf_cutoff)
saveRDS(ligand_target_matrix,"Results/ligand_target_matrixWithweights.rds")
saveRDS(weighted_networks,"Results/weighted_networksWithSourceWeights.rds")
```

